---
title: "Bulkseq_scseq_comparison_d1"
author: "Khang"
date: "`r Sys.Date()`"
output:
  pdf_document: 
    keep_tex: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,message = FALSE,warning = FALSE)
knitr::opts_chunk$set(dev = "pdf",
                      dpi = 600, out.width = "100%", out.height = "100%",
                      cache = TRUE)
```

```{r libs, include=FALSE}
library(immunarch)
library(tidyverse)
library(pheatmap)
library(ggplotify)
library(cowplot)
library(reshape2)
library(seqinr)
library(RColorBrewer)
library(cluster)
library(rstatix)
library(ggpubr)
library(vegan)
library(scales)
```

```{r eval=FALSE, include=FALSE}
# Load the data
data_d1_bulk <- readRDS("./data_d1_bulkseq.RDS")
data_d1_sc <- readRDS("./data_d1_scseq.RDS")

# Filter the data to keep only heavy chain and remove IgE samples from bulkseq (very rare & not present in scseq)
data_d1_bulk$data <- data_d1_bulk$data[str_which(names(data_d1_bulk$data),"IgA|IgD|IgG|IgM")]
data_d1_sc$data <- data_d1_sc$data[str_which(names(data_d1_sc$data),"IgA|IgD|IgG|IgM")]

# Merge the bulk and sc data into one object
data_d1_both <- list(data = c(data_d1_bulk$data, data_d1_sc$data),
                      meta = rbind(data_d1_bulk$meta, data_d1_sc$meta))

# Format the metadata
data_d1_both$meta <- tibble(Sample = names(data_d1_both$data),
                            Sample_Id = str_split(names(data_d1_both$data), "_",simplify = T,)[,2],
                            Accumulation = str_count(str_split(names(data_d1_both$data), "_",simplify = T,)[,2], "-") + 1,
                            Isotype = str_split(names(data_d1_both$data), "_",simplify = T,)[,3],
                            Method = str_split(names(data_d1_both$data), "_",simplify = T,)[,4],
                            Chain = "hc")
                                
# Save the data
saveRDS(data_d1_both, "./data_d1_both.RDS")
```

```{r}
# Load the data
data_d1_both <- readRDS("./data_d1_both.RDS")

#Replace the method sc with single-cell
data_d1_both$meta$Method <- str_replace_all(data_d1_both$meta$Method, "sc", "single-cell")
```

```{r}
# Create an annotation df for coloring
# Create a separate df with rownames matching the sample names
ann_df <- data.frame(#Sample_Id = data_d1_both$meta$Sample_Id,
                     Isotype = data_d1_both$meta$Isotype,
                     Accumulation = data_d1_both$meta$Accumulation,
                     Method = data_d1_both$meta$Method)
rownames(ann_df) <- names(data_d1_both$data)

# Specify colors for the annotation
ann_color <- list(#Sample_Id = brewer.pal(6,"Set3"),
                  #Isotype = c("#2894BF","#395896","#41868e","#7caf67","#824f7e"),
                  Isotype = brewer.pal(4,"Set2"),
                  Accumulation = brewer.pal(8,"Purples"),
                  Method = c("#404040","#bababa"))

#names(ann_color$Sample_Id) <- unique(data_d1_both$meta$Sample_Id)
names(ann_color$Isotype) <- unique(data_d1_both$meta$Isotype)
names(ann_color$Accumulation) <- unique(data_d1_both$meta$Accumulation)
names(ann_color$Method) <- unique(data_d1_both$meta$Method)
```

# Unique CDR3 count
```{r}
cdr3_count <- tibble(Sample = names(data_d1_both$data),
                     Volume = unlist(map(data_d1_both$data, function(x) length(unique(x$CDR3.aa)))))
cdr3_count <- merge(cdr3_count, data_d1_both$meta)
cdr3_count <- cdr3_count %>% filter(Accumulation == 1)
cdr3_count$Isotype <- factor(cdr3_count$Isotype, levels = c("IgD", "IgM", "IgA", "IgG"))

# Reorder the factors by isotype
cdr3_count <- cdr3_count %>% arrange(Isotype) %>% mutate(Sample = factor(Sample, levels = Sample))
```

```{r fig.height=4, fig.width=14}
p_cdr3_count_bulk <- ggplot(data = cdr3_count %>% filter(Method == "bulk"), aes(x = Sample, y = Volume, fill = Isotype)) +
  geom_col() +
  scale_fill_manual(values = c("#FC8D62", "#E78AC3", "#66C2A5", "#8DA0CB")) +
  geom_text(aes(label = comma(Volume)), size = 2.5, vjust = -0.2) +
  scale_y_continuous(labels = comma) +
  theme_half_open() +
  theme(axis.text.x = element_text(angle = 90), legend.position = "none") +
  labs(x = "Sample", y = "Unique CDRH3 count")

p_cdr3_count_sc <- ggplot(data = cdr3_count %>% filter(Method == "single-cell"), aes(x = Sample, y = Volume, fill = Isotype)) +
  geom_col() +
  scale_fill_manual(values = c("#FC8D62", "#E78AC3", "#66C2A5", "#8DA0CB")) +
  geom_text(aes(label = comma(Volume)), size = 2.5, vjust = -0.2) +  
  scale_y_continuous(labels = comma) +
  theme_half_open() +
  theme(axis.text.x = element_text(angle = 90), legend.position = "none") +
  labs(x = "Sample", y = "Unique CDRH3 count")

p_cdr3_count <- plot_grid(p_cdr3_count_bulk, p_cdr3_count_sc, rel_widths = c(1,5))
p_cdr3_count

#ggsave("./figures/cdrh3_count_d1.pdf", p_cdr3_count, width = 14, height = 4, dpi = 600)
#ggsave("./figures/cdrh3_count_d1.png", p_cdr3_count, width = 14, height = 4, dpi = 600)
```

# Evenness profile
```{r}
# Calculate the Hill numbers
hill_all <- lapply(data_d1_both$data, function (x) renyi(x$Proportion, scales = seq(0, 10, by = 0.2), hill = TRUE) %>% as_tibble(.,rownames = "alpha"))

# Convert diversity into Evenness
evenness_all <- sapply(hill_all, function (x) x$value <- x$value/x$value[[1]])
evenness_all <- as_tibble(evenness_all)

# Reshape the data from wide to long format
evenness_all_long <- gather(evenness_all, key = "Sample", value = "Value")
Alpha <- seq(0, 10, by = 0.2)
evenness_all_long <- cbind(evenness_all_long, Alpha)
evenness_all_long <- merge(evenness_all_long, data_d1_both$meta) %>% filter(Accumulation == 1)
```

```{r eval=FALSE, fig.height=3, fig.width=7, include=FALSE}
p_evenness_bulk <- ggplot(evenness_all_long %>% filter(Method == "bulk"), aes(x = Alpha, y = Value, group = Sample)) +
  geom_line(aes(color = Isotype), size = 0.5) +
  scale_color_brewer(type = "qual", palette = "Set2") +
  theme_half_open() +
  theme(panel.grid.major = element_line(color = "gray80", size = 0.2, linetype = 1), legend.position = "none") + 
  labs(x = "Alpha", y = "Repertoire evenness") +
  scale_x_continuous(breaks = seq(0, 10, by = 1)) + #change the break of the tick marks
  scale_y_continuous(breaks = seq(0.0, 1.0, by = 0.1))

p_evenness_sc <- ggplot(evenness_all_long %>% filter(Method == "single-cell"), aes(x = Alpha, y = Value, group = Sample)) +
  geom_line(aes(color = Isotype), size = 0.5) +
  scale_color_brewer(type = "qual", palette = "Set2") +
  theme_half_open() +
  theme(panel.grid.major = element_line(color = "gray80", size = 0.2, linetype = 1),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(), axis.line.y = element_blank()) + # Remove y-axis
  labs(x = "Alpha", y = "") +
  scale_x_continuous(breaks = seq(0, 10, by = 1)) + #change the break of the tick marks
  scale_y_continuous(breaks = seq(0.0, 1.0, by = 0.1))

p_evenness <- plot_grid(p_evenness_bulk, p_evenness_sc, rel_widths = c(1,1.2))
p_evenness

#ggsave("./figures/evenness_d1.pdf", p_evenness, width = 7, height = 3, dpi = 600)
#ggsave("./figures/evenness_d1.png", p_evenness, width = 7, height = 3, dpi = 600)
```

```{r fig.height=3.5, fig.width=7}
p_evenness <- ggplot(evenness_all_long, aes(x = Alpha, y = Value, group = Sample)) +
  geom_line(aes(color = Isotype), size = 0.5) +
  scale_color_brewer(type = "qual", palette = "Set2") +
  theme_half_open() +
  theme(panel.grid.major = element_line(color = "gray80", size = 0.2, linetype = 1), legend.position = "right", strip.background = element_blank()) + 
  labs(x = "Alpha", y = "Repertoire evenness") +
  scale_x_continuous(breaks = seq(0, 10, by = 1)) + #change the break of the tick marks
  scale_y_continuous(breaks = seq(0.0, 1.0, by = 0.1)) +
  facet_wrap(vars(Method))

p_evenness

#ggsave("./figures/evenness_d1.pdf", p_evenness, width = 7, height = 3.5, dpi = 600)
#ggsave("./figures/evenness_d1.png", p_evenness, width = 7, height = 3.5, dpi = 600)
```


# VH gene usage
```{r VH-gene_usage, message=FALSE, warning=FALSE}
# V gene usage heavy chain (don't weigh by clone size)
gene_use_hv <- geneUsage(data_d1_both$data, .gene = c("hs.ighv"), .ambig = "maj", .norm = TRUE, .quant = NA)
gene_use_hv[is.na(gene_use_hv)] <- 0

gene_use_hv_bulk <- geneUsage(data_d1_both$data[str_which(data_d1_both$meta$Method, "bulk")], .gene = c("hs.ighv"), .ambig = "maj", .norm = TRUE, .quant = NA)
gene_use_hv_bulk[is.na(gene_use_hv_bulk)] <- 0

gene_use_hv_sc <- geneUsage(data_d1_both$data[intersect(str_which(data_d1_both$meta$Method, "single-cell"), str_which(data_d1_both$meta$Accumulation , "1"))], .gene = c("hs.ighv"), .ambig = "maj", .norm = TRUE, .quant = NA)
gene_use_hv_sc[is.na(gene_use_hv_sc)] <- 0
```

```{r}
# Calculate mean and sd of VH-gene usage
mean_gene_use_hv_bulk <- tibble(Names = gene_use_hv_bulk$Names,
                                Mean = apply(gene_use_hv_bulk[,-1], 1, mean),
                                Sd = apply(gene_use_hv_bulk[,-1], 1, sd))

mean_gene_use_hv_sc <- tibble(Names = gene_use_hv_sc$Names,
                                Mean = apply(gene_use_hv_sc[,-1], 1, mean),
                                Sd = apply(gene_use_hv_sc[,-1], 1, sd))
```

```{r fig.height=5, fig.width=18}
p_mean_gene_use_hv_bulk <- ggplot(data = mean_gene_use_hv_bulk, aes(x = reorder(Names, Mean, decreasing = T), y = Mean)) +
  geom_col(fill = "#588DE0") +
  geom_errorbar(aes(ymin = Mean - Sd, ymax = Mean + Sd), width = 0.4, alpha = 0.8) +
  scale_y_continuous(breaks = seq(0, 0.15, by = 0.05)) +
  theme_half_open() +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(x = "V-gene name", y = "V-gene frequency")

p_mean_gene_use_hv_sc <- ggplot(data = mean_gene_use_hv_sc, aes(x = reorder(Names, Mean, decreasing = T), y = Mean)) +
  geom_col(fill = "#588DE0") +
  geom_errorbar(aes(ymin = Mean - Sd, ymax = Mean + Sd), width = 0.4, alpha = 0.8) +
  scale_y_continuous(breaks = seq(0, 0.15, by = 0.05)) +
  theme_half_open() +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(x = "V-gene name", y = "V-gene frequency")

p_mean_gene_use_hv <- plot_grid(p_mean_gene_use_hv_bulk, p_mean_gene_use_hv_sc, rel_widths = c(1,1.25))
p_mean_gene_use_hv

#ggsave("./figures/gene_use_hv_d1.pdf", p_mean_gene_use_hv, width = 18, height = 5, dpi = 600)
#ggsave("./figures/gene_use_hv_d1.png", p_mean_gene_use_hv, width = 18, height = 5, dpi = 600)
```

# Display all VH-gene usage as faceted barplots
```{r}
# Transform the V gene usage matrix into long format
gene_use_hv_long <- column_to_rownames(gene_use_hv, var = "Names")
gene_use_hv_long <- pivot_longer(as_tibble(gene_use_hv_long), cols = everything(), cols_vary = "slowest", names_to = "Sample", values_to = "Frequency")
gene_use_hv_long$Gene <- rep(gene_use_hv$Names, times = ncol(gene_use_hv)-1)
gene_use_hv_long <- merge(gene_use_hv_long, data_d1_both$meta)
gene_use_hv_long <- gene_use_hv_long %>% filter(Accumulation == 1) # exclude the merged replicates sample

# Calculate mean and sd of VH-gene usage
#group by gene, isotype, method, calculate the mean and sd for the frequency of each VH gene into a summary tibble

gene_use_hv_long_summary <- gene_use_hv_long %>% group_by(Method, Isotype, Gene) %>% summarise(Mean_freq = mean(Frequency), Sd_freq = sd(Frequency))

# Count the number of genes present in each facet
gene_use_hv_long_summary %>% group_by(Method, Isotype) %>% summarise(Gene_count = sum(Mean_freq > 0))
```

```{r fig.height=6, fig.width=14}
p_gene_use_hv_facet <- ggplot(data = gene_use_hv_long_summary, aes(x = Gene, y = Mean_freq)) +
  geom_col(aes(fill = Isotype)) +
  scale_fill_brewer(type = "qual", palette = "Set2") +
  geom_errorbar(aes(ymin = Mean_freq - Sd_freq, ymax = Mean_freq + Sd_freq), width = 0.3, alpha = 0.5) +
  geom_text(aes(label = round(Mean_freq, 2)), vjust = -0.2, size = 1) +
  scale_y_continuous(breaks = seq(0, 0.15, by = 0.05)) +
  theme_half_open() +
  theme(axis.text.x = element_text(angle = 90, size = 7), strip.background = element_blank(), text = element_text(size = 20)) +
  facet_grid(rows = vars(Isotype), cols = vars(Method)) +
  labs(x = "V-gene name", y = "V-gene frequency")

p_gene_use_hv_facet

#ggsave("./figures/gene_use_hv_facet_d1.pdf", p_gene_use_hv_facet, width = 14, height = 6, dpi = 600)
```


# Display all VH-gene usage as heatmap (unused)
```{r eval=FALSE, include=FALSE}
gene_use_hv2 <- geneUsage(data_d1_both$data[str_which(data_d1_both$meta$Accumulation, "1")], .gene = c("hs.ighv"), .ambig = "maj", .norm = TRUE, .quant = NA)
gene_use_hv2_mat <- t(column_to_rownames(gene_use_hv2, var = "Names"))
```

```{r eval=FALSE, fig.height=6, fig.width=12, include=FALSE}
p_gene_use_hv_hm <- pheatmap(gene_use_hv2_mat, cluster_rows = F, cluster_cols = F, na_col = "white", border_color = "white",
         angle_col = 90, annotation_row = ann_df[str_which(data_d1_both$meta$Accumulation, "1"), -2], annotation_colors = ann_color[-2],
         color = colorRampPalette(c("#c6dbef", "#084594"))(length(seq(round(min(gene_use_hv2_mat, na.rm = T),1), round(max(gene_use_hv2_mat, na.rm = T),1), by = 0.01))),
         legend_breaks = seq(round(min(gene_use_hv2_mat, na.rm = T),1), round(max(gene_use_hv2_mat, na.rm = T),1), by = 0.02),
         legend_labels = c(seq(round(min(gene_use_hv2_mat, na.rm = T),1), round(max(gene_use_hv2_mat, na.rm = T),1) - 0.02, by = 0.02), "Frequency"), # The breaks vector needs to be ONE element longer than the color vector
)

#ggsave("./figures/gene_use_hv_hm_d1.pdf", p_gene_use_hv_hm, width = 12, height = 6, dpi = 600)
#ggsave("./figures/gene_use_hv_hm_d1.png", p_gene_use_hv_hm, width = 12, height = 6, dpi = 600)
```


# Compare the number of unique VH genes in bulkseq samples and in scseq from 1 to 8 merged
```{r fig.height=6, fig.width=8}
# Count the number of VH gene detected in each sample (freq > 0)
gene_count_hv <- tibble(Sample = colnames(gene_use_hv[-1]),
                        Gene_count = unlist(apply(gene_use_hv[,-1], MARGIN = 2, function(x) length(which(x > 0)))))

gene_count_hv <- merge(gene_count_hv, data_d1_both$meta)

#gene_count_hv <- gene_count_hv[order(gene_count_hv$Accumulation),]
```


```{r fig.height=6, fig.width=8}
p_gene_count_hv <- ggplot(data = gene_count_hv %>% filter(Method == "single-cell"), 
       mapping = aes(x = reorder(Sample, Gene_count) , y = Gene_count)) +
  geom_col(aes(fill = as.character(Accumulation)), position= "dodge") +
  scale_fill_manual(values = brewer.pal(9, "Blues")[-1]) +
  theme_half_open() +
  theme(legend.position = "bottom", axis.text.x = element_blank()) +
  labs(#title = "(Dataset 1) Number of unique VH genes detected in ScBCR-seq",
       x = "Sample", y = "Unique VH gene count", fill = "Number of ScBCR-seq replicate merged")

p_gene_count_hv

#ggsave("./figures/gene_count_hv.pdf", p_gene_count_hv, width = 8, height = 6, dpi = 600)
#ggsave("./figures/gene_count_hv.png", p_gene_count_hv, width = 8, height = 6, dpi = 600)
```

# VH gene usage correlation

```{r VH-gene_usage_correlation_isotype}
# Calculate pairwise correlation metrics and JSD
vh_cor_pearson <- geneUsageAnalysis(gene_use_hv, .method = "cor", .cor = "pearson", .verbose = F)

# Calculate the distance based on correlation values
dist_pearson <- as.dist(1-cor(gene_use_hv[,-1], method = "pearson", use = "na.or.complete"))
hc_pearson <- hclust(dist_pearson, method = "ward.D2")

```


```{r VH-gene_usage_correlation_isotype, fig.height=14, fig.width=14}
# Draw the cor plot (remember to make silhouette plot to justify cluster number)
p_cor_pearson <- as.ggplot(pheatmap(vh_cor_pearson, cluster_rows = T, cluster_cols = T, na_col = "gray80", silent = T, clustering_method = "ward.D2",
                     clustering_distance_rows = dist_pearson, clustering_distance_cols = dist_pearson,
                     border_color = "white", angle_col = 90, fontsize = 14,
                     show_colnames = F, show_rownames = F, annotation_names_row = F,
                     display_numbers = F, number_color = "black",
                     annotation = ann_df, annotation_row = ann_df,
                     annotation_colors = ann_color,
                     cutree_rows = 4, cutree_cols = 4,
                     breaks = seq(round(min(vh_cor_pearson, na.rm = T),1), 1.0, by = 0.05),
                     legend_breaks = seq(round(min(vh_cor_pearson, na.rm = T),1), 1.0, by = 0.05),
                     legend_labels = c(seq(round(min(vh_cor_pearson, na.rm = T),1), 0.95, by = 0.05), "Cor"), # The breaks vector needs to be ONE element longer than the color vector
                     color = colorRampPalette(c("#ffffb2", "#e31a1c"))(length(seq(round(min(vh_cor_pearson, na.rm = T),1), 1.0, by = 0.05))),
                     ))

p_cor_pearson

#ggsave("./figures/VH_Pearson_cor_heatmap_d1.pdf", p_cor_pearson, width = 14, height = 14, dpi = 600)
```

```{r}
# Visualize the pearson correlation of VH gene usage using box violin plots

# Set the upper triangle to pairwise matrix to NA
vh_cor_pearson2 <- vh_cor_pearson
vh_cor_pearson2[lower.tri(vh_cor_pearson2)] <- NA

# Convert matrix to long format
vh_cor_pearson_long <- pivot_longer(as_tibble(vh_cor_pearson2), cols = everything(), names_to = "Target", values_to = "Cor")
vh_cor_pearson_long$Sample <- rep(rownames(vh_cor_pearson2), each = ncol(vh_cor_pearson2))
vh_cor_pearson_long <- merge(vh_cor_pearson_long, data_d1_both$meta)
vh_cor_pearson_long$Sample_Id_Target <- str_split(vh_cor_pearson_long$Target, "_", simplify = T)[,2]
vh_cor_pearson_long$Accumulation_Target <- str_count(str_split(vh_cor_pearson_long$Target, "_", simplify = T)[,2], "-") + 1
vh_cor_pearson_long$Isotype_Target <- str_split(vh_cor_pearson_long$Target, "_", simplify = T)[,3]
vh_cor_pearson_long$Method_Target <- str_split(vh_cor_pearson_long$Target, "_", simplify = T)[,4]
vh_cor_pearson_long$Method_Target <- str_replace_all(vh_cor_pearson_long$Method_Target, "sc", "single-cell")

# Set up groups to compare between bulk and scseq, group by number of accumulation in scseq
vh_cor_pearson_long$Same_Isotype <- ifelse(vh_cor_pearson_long$Isotype == vh_cor_pearson_long$Isotype_Target, "Same Isotype", "Different Isotypes")
vh_cor_pearson_long$Same_Method <- ifelse(vh_cor_pearson_long$Method == vh_cor_pearson_long$Method_Target, "Same method", "Different methods")


vh_cor_pearson_long$Pairwise_comparison <- as.character(vh_cor_pearson_long$Accumulation_Target)

# Remove duplicates and self-comparison by removing NA from the df
vh_cor_pearson_long <- vh_cor_pearson_long %>% filter(!is.na(Cor))
```


```{r}
# Filter for only comparison btw bulk & sc samples
vh_cor_pearson_long_bvs <- vh_cor_pearson_long %>% filter(Method == "bulk" & Method_Target == "single-cell")

# Stat test for difference within a group (group by pwc)
cor_kruskal <- vh_cor_pearson_long_bvs %>% kruskal_test(Cor~Pairwise_comparison)

# Calculate the effect size (how much pwc group contribute to variation in cor)
cor_kruskal_eff <- vh_cor_pearson_long_bvs %>% kruskal_effsize(Cor~Pairwise_comparison)
cor_kruskal_eff

# In the group with difference, calculate pairwise mean comparison with multiple testing correction
cor_pwc <- vh_cor_pearson_long_bvs %>% wilcox_test(Cor~Pairwise_comparison, p.adjust.method = "bonferroni")

# Auto add coordinates for plotting
cor_pwc <- cor_pwc %>% add_xy_position(fun = "max", x = "Pairwise_comparison")

cor_pwc_median <- vh_cor_pearson_long_bvs %>% group_by(Pairwise_comparison) %>% summarise(median = median(Cor))

# Add a line for median Cor of bvb and a line for median cor of svs
cor_pwc_median_bvb <- vh_cor_pearson_long %>% filter(Same_Method == "Same method" & Method == "bulk") %>% summarise(median = median(Cor))
cor_pwc_median_svs <- vh_cor_pearson_long %>% filter(Same_Method == "Same method" & Method == "single-cell" & Accumulation == 1 & Accumulation_Target == 1) %>% summarise(median = median(Cor))
```

```{r fig.height=4, fig.width=5.5}
p_cor_pearson_box <- ggplot(data = vh_cor_pearson_long_bvs, aes(x = Pairwise_comparison, y = Cor)) +
  geom_violin(fill = "#588DE0") +
  #scale_fill_brewer(type = "seq", palette = "Purples") +
  geom_boxplot(fill = "#588DE0", width = 0.1, color= "black") +
  geom_hline(color = "#2166ac", linetype = 2, yintercept = cor_pwc_median_bvb$median) +
  geom_hline(color = "#b2182b", linetype = 2, yintercept = cor_pwc_median_svs$median) +
  geom_text(x = 7.5, y = 0.97, label = "bulk vs bulk", check_overlap = T) +
  geom_text(x = 7.5, y = 0.89, label = "sc vs sc", check_overlap = T) +
  scale_y_continuous(breaks = c(0.30, 0.40, 0.50, 0.60, 0.70, 0.80, cor_pwc_median_svs$median, 0.90, cor_pwc_median_bvb$median, 1.00),
                     labels = c(0.30, 0.40, 0.50, 0.60, 0.70, 0.80, round(cor_pwc_median_svs$median,2), 0.90, round(cor_pwc_median_bvb$median,2), 1.00)) +
  geom_text(data = cor_pwc_median, aes(x = Pairwise_comparison, y = 0.45, label = round(median,2))) +
  stat_compare_means(method = "kruskal.test", label.y = 1) +
  stat_pvalue_manual(data = cor_pwc, hide.ns = TRUE,  step.increase = 0.05) +
  theme_half_open() +
  theme(legend.position = "none") +
  labs(x = "Number of scBCR-seq replicates merged", y = "VH-gene usage Pearson correlation")

p_cor_pearson_box

#ggsave("./figures/VH_Pearson_cor_accumulation_d1.pdf", p_cor_pearson_box, width = 5.5, height = 4, dpi = 600)
#ggsave("./figures/VH_Pearson_cor_accumulation_d1.png", p_cor_pearson_box, width = 5.5, height = 4, dpi = 600)
```

\newpage

# Repertoire overlap (shared CDRH3 sequence)

```{r}
cdrh3_overlap_jaccard <- (repOverlap(data_d1_both$data, .method = "jaccard",.col = "aa", .verbose = F))

cdrh3_overlap_jaccard_bvs <- cdrh3_overlap_jaccard[str_which(colnames(cdrh3_overlap_jaccard), "sc"),str_which(rownames(cdrh3_overlap_jaccard), "bulk")]
```

```{r Repertoire_overlap_isotype_jaccard, fig.height=14, fig.width=14}
p_cdrh3_overlap_jaccard_isotype <- as.ggplot(
  pheatmap(cdrh3_overlap_jaccard, cluster_rows = T, cluster_cols = T, clustering_method = "ward.D2",
           cutree_rows = 5, cutree_cols = 5,
           border_color = "white", #na_col =  "gray3",
           show_colnames = F, show_rownames = F, annotation_names_row = F,
           angle_col = 90, display_numbers = F, number_color = "black",
           number_format = "%.2f", silent = T, fontsize_number = 10,
           annotation = ann_df, annotation_row = ann_df,
           annotation_colors = ann_color,
           breaks = seq(round(min(cdrh3_overlap_jaccard, na.rm = T),1), 1.0, by = 0.05),
           legend_breaks = seq(round(min(cdrh3_overlap_jaccard, na.rm = T),1), 1.0, by = 0.05),
           legend_labels = c(seq(round(min(cdrh3_overlap_jaccard, na.rm = T),1), 0.95, by = 0.05), "Overlap"), # The breaks vector needs to be ONE element longer than the color vector
           color = colorRampPalette(c("#ffffb2", "#e31a1c"))(20)
           ))

p_cdrh3_overlap_jaccard_isotype

#ggsave("./figures/Repertoire_overlap_jaccard_d1.pdf", p_cdrh3_overlap_jaccard_isotype, width = 14, height = 14, dpi = 600)
```

```{r fig.height=16, fig.width=8}
p_cdrh3_overlap_jaccard_bvs <- as.ggplot(
  pheatmap(cdrh3_overlap_jaccard_bvs, cluster_rows = T, cluster_cols = F, clustering_method = "ward.D2",
           #cutree_rows = 3, cutree_cols = 3,
           border_color = "white", #na_col =  "gray3",
           main = "CDRH3 Jaccard overlap bulkBCR-seq vs scBCR-seq",
           angle_col = 90, display_numbers = T, number_color = "black",
           number_format = "%.4f", silent = T, fontsize_number = 10,
           annotation = ann_df, annotation_row = ann_df,
           annotation_colors = ann_color,
           color = colorRampPalette(c("#ffffb2", "#e31a1c"))(25)
           ))

p_cdrh3_overlap_jaccard_bvs
```

# Quantify the difference in Jaccard overlap for bvs between 1-8 merged samples
```{r}
# Convert matrix to long format
cdrh3_overlap_jaccard_bvs_long <- pivot_longer(as_tibble(t(cdrh3_overlap_jaccard_bvs)), cols = everything(), names_to = "Target", values_to = "Overlap")
cdrh3_overlap_jaccard_bvs_long$Sample <- rep(colnames(cdrh3_overlap_jaccard_bvs), each = nrow(cdrh3_overlap_jaccard_bvs))
cdrh3_overlap_jaccard_bvs_long <- merge(cdrh3_overlap_jaccard_bvs_long, data_d1_both$meta)
cdrh3_overlap_jaccard_bvs_long$Sample_Id_Target <- str_split(cdrh3_overlap_jaccard_bvs_long$Target, "_", simplify = T)[,2]
cdrh3_overlap_jaccard_bvs_long$Accumulation_Target <- str_count(str_split(cdrh3_overlap_jaccard_bvs_long$Target, "_", simplify = T)[,2], "-") + 1
cdrh3_overlap_jaccard_bvs_long$Isotype_Target <- str_split(cdrh3_overlap_jaccard_bvs_long$Target, "_", simplify = T)[,3]
cdrh3_overlap_jaccard_bvs_long$Method_Target <- str_split(cdrh3_overlap_jaccard_bvs_long$Target, "_", simplify = T)[,4]
cdrh3_overlap_jaccard_bvs_long$Method_Target <- str_replace_all(cdrh3_overlap_jaccard_bvs_long$Method_Target, "sc", "single-cell")

# Set up groups to compare between bulk and scseq, group by number of accumulation in scseq
cdrh3_overlap_jaccard_bvs_long$Same_Isotype <- ifelse(cdrh3_overlap_jaccard_bvs_long$Isotype == cdrh3_overlap_jaccard_bvs_long$Isotype_Target, "Same isotype", "Different isotypes")
cdrh3_overlap_jaccard_bvs_long$Same_Method <- ifelse(cdrh3_overlap_jaccard_bvs_long$Method == cdrh3_overlap_jaccard_bvs_long$Method_Target, "Same method", "Different methods")
cdrh3_overlap_jaccard_bvs_long$Pairwise_comparison <- as.character(cdrh3_overlap_jaccard_bvs_long$Accumulation_Target)
```

```{r}
# Calculate pairwise mean comparison between same vs diff isotypes
overlap_pwc_bvs_di <- cdrh3_overlap_jaccard_bvs_long %>% wilcox_test(Overlap~Same_Isotype, p.adjust.method = "bonferroni")

# Auto add coordinates for plotting
overlap_pwc_bvs_di <- overlap_pwc_bvs_di %>% add_xy_position(fun = "max", x = "Same_Isotype")

overlap_pwc_median_bvs_di <- cdrh3_overlap_jaccard_bvs_long %>% group_by(Same_Isotype) %>% summarise(median = median(Overlap))
```

```{r fig.height=4, fig.width=3, warning=FALSE}
p_overlap_box_bvs_di <- ggplot(data = cdrh3_overlap_jaccard_bvs_long, aes(x = Same_Isotype, y = Overlap)) +
  geom_violin(fill = "#588DE0") +
  #scale_fill_brewer(type = "qual", palette = 1, direction = -1) +
  scale_y_log10(breaks = c(1e-05,1e-04, 1e-03, 1e-02, 1e-01, 1e+00)) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
  geom_boxplot(fill = "#588DE0", width = 0.075, color= "black") +
  geom_text(data = overlap_pwc_median_bvs_di, aes(x = Same_Isotype, y = 1e-06, label = round(median,5))) +
  stat_pvalue_manual(data = overlap_pwc_bvs_di, hide.ns = TRUE,  step.increase = 0.005) +
  theme_half_open() +
  theme(legend.position = "none") +
  labs(x = "Isotype comparison", y = "CDRH3 Jaccard overlap")

p_overlap_box_bvs_di

#ggsave("./figures/Jaccard_overlap_box_bvs_di_d1.pdf", p_overlap_box_bvs_di, width = 3, height = 4, dpi = 600)
#ggsave("./figures/Jaccard_overlap_box_bvs_di_d1.png", p_overlap_box_bvs_di, width = 3, height = 4, dpi = 600)
```

```{r}
# Restrict the bvs comparison to only from same isotype (bc diff isotype has Jaccard overlap mostly close to 0)
cdrh3_overlap_jaccard_bvs_long_si <- cdrh3_overlap_jaccard_bvs_long %>% filter(Same_Isotype == "Same isotype")

# Stat test for difference within a group (group by pwc)
overlap_kruskal_bvs_si <- cdrh3_overlap_jaccard_bvs_long_si %>% kruskal_test(Overlap~Pairwise_comparison)

# Calculate the effect size (how much pwc group contribute to variation in overlap)
overlap_kruskal_eff_bvs_si <- cdrh3_overlap_jaccard_bvs_long_si %>% kruskal_effsize(Overlap~Pairwise_comparison)
overlap_kruskal_eff_bvs_si

# In the group with difference, calculate pairwise mean comparison with multiple testing correction
overlap_pwc_bvs_si <- cdrh3_overlap_jaccard_bvs_long_si %>% wilcox_test(Overlap~Pairwise_comparison, p.adjust.method = "bonferroni")

# Auto add coordinates for plotting
overlap_pwc_bvs_si <- overlap_pwc_bvs_si %>% add_y_position(fun = "max", scales = "free_y")

overlap_pwc_median_bvs_si <- cdrh3_overlap_jaccard_bvs_long_si %>% group_by(Pairwise_comparison) %>% summarise(median = median(Overlap))
```

```{r fig.height=4, fig.width=6.5, warning=FALSE}
p_overlap_box_bvs_si <- ggplot(data = cdrh3_overlap_jaccard_bvs_long_si, aes(x = Pairwise_comparison, y = Overlap)) +
  geom_violin(fill = "#588DE0") +
  #scale_fill_brewer(type = "seq", palette = "Greens") +
  geom_boxplot(fill = "#588DE0", width = 0.1, color= "black") +
  geom_text(data = overlap_pwc_median_bvs_si, aes(x = Pairwise_comparison, y = -0.002, label = round(median,4))) +
  stat_compare_means(method = "kruskal.test", label.y = 0.02) +
  stat_pvalue_manual(data = overlap_pwc_bvs_si, hide.ns = TRUE, label = "p.adj", label.size = 2.5, step.increase = 0.005) +
  #scale_y_log10() +
  theme_half_open() +
  theme(legend.position = "none") +
  labs(x = "Number of scBCR-seq replicates merged", y = "Same isotype CDRH3 Jaccard overlap")

p_overlap_box_bvs_si

#ggsave("./figures/Jaccard_overlap_box_bvs_si_d1.pdf", p_overlap_box_bvs_si, width = 6.5, height = 4, dpi = 600)
#ggsave("./figures/Jaccard_overlap_box_bvs_si_d1.png", p_overlap_box_bvs_si, width = 6.5, height = 4, dpi = 600)
```